Estimating the association between Facebook adoption and well-being in 72 countries

### 2022-11-09 data-prepare.Rmd
Commit synthetic and population data

@@

```{r}
library(readxl)
pop <- read_excel(
  "data/WPP2022_POP_F01_1_POPULATION_SINGLE_AGE_BOTH_SEXES.xlsx", 
  sheet = 1, 
  skip = 16,
  col_types = "text",
  .name_repair = make_clean_names
  )
pop <- pop %>% 
  drop_na(iso3_alpha_code) %>% 
  select(
    iso3_alpha_code, year, starts_with("x")
  ) %>% 
  mutate(
    country = countrycode(
      iso3_alpha_code, 
      origin = "iso3c", 
      destination = "cldr.short.en"
    )
  )
pop <- pop %>% 
  filter(
    country %in% fb_countries,
    between(year, 2008, 2019)
  ) %>% 
  pivot_longer(
    starts_with("x"), 
    names_to = "age",
    values_to = "pop",
    names_transform = ~str_remove(., "x") %>% as.numeric, 
    values_transform = ~ as.numeric(.) * 1000
  ) %>% 
  filter(between(age, 13, 100)) %>% 
  mutate(
    age = cut(age, c(0, 34.5, 1000), labels = c("13-34", "35+"))
  )
pop <- pop %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(country, year, age) %>% 
  summarise(pop = sum(pop)) %>% 
  ungroup()
```

@@

```{r gallup-save}
# Save
dir.create("data", FALSE)
write_rds(gwp, "data/gwp.rds")
write_rds(pop, "data/population.rds")
```

>>>>

## GBD

```{r}
# Load data files and merge to one table
gbd <- list.files(
  here("data-raw/GBD/"),
  pattern = ".zip", full.names = TRUE, recursive = TRUE
) %>%
  read_csv()

# First step is to clean names and remove unnecessary variables
gbd <- clean_names(gbd)
# distinct(gbd, measure, metric)
gbd <- select(gbd, -measure, -metric)

# Clean cause names
# distinct(gbd, cause)
gbd <- gbd %>%
  mutate(
    cause = case_when(
      cause == "Anxiety disorders" ~ "Anxiety",
      cause == "Depressive disorders" ~ "Depression",
      cause == "Self-harm" ~ "Selfharm"
    )
  )

# Harmonise variable names with other datasets
gbd <- gbd %>%
  rename(country = location, outcome = cause)

# The outcomes values are model predictions and come with lower and upper CI limits (2.5 and 97.5 %iles of their posterior distributions). We convert those to normal approximate standard errors.
gbd <- gbd %>%
  mutate(se = (upper - lower) / (1.96 * 2)) %>%
  select(-c(upper, lower))

# Harmonise old names and replace only if harmonised name found
gbd <- gbd %>%
  mutate(
    country_harmonised = countryname(
      country,
      destination = "cldr.short.en"
    )
  ) %>%
  # This picks the harmonised name if exists, otherwise original name
  mutate(country = coalesce(country_harmonised, country)) %>%
  select(-country_harmonised)

# Include only FB years and countries
gbd <- gbd %>% 
  filter(country %in% fb_countries) %>% 
  filter(between(year, 2008, 2019))

# Categorize ages
gbd <- gbd %>%
  mutate(
    age = factor(
      if_else(
        age %in% c("10 to 14", "15 to 19", "20 to 24", "25 to 29", "30 to 34"), 
        "13-34", 
        "35+"
      )
    )
  )
```

```{r}
# Aggregate to new age groups
gbd <- gbd %>% 
  group_by(country, year, sex, age, outcome) %>%
  summarise(
    val = mean(val, na.rm = TRUE),
    se = sqrt(mean(se^2))  # average standard error
  ) %>%
  ungroup()
```

@@

```{r}
library(readxl)
pop <- read_excel(
  "data/WPP2022_POP_F01_1_POPULATION_SINGLE_AGE_BOTH_SEXES.xlsx", 
  sheet = 1, 
  skip = 16,
  col_types = "text",
  .name_repair = make_clean_names
)
pop <- pop %>% 
  drop_na(iso3_alpha_code) %>% 
  select(
    iso3_alpha_code, year, starts_with("x")
  ) %>% 
  mutate(
    country = countrycode(
      iso3_alpha_code, 
      origin = "iso3c", 
      destination = "cldr.short.en"
    ),
    year = as.numeric(year)
  )
pop <- pop %>% 
  filter(
    country %in% fb_countries,
    between(year, 2008, 2019)
  ) %>% 
  pivot_longer(
    starts_with("x"), 
    names_to = "age",
    values_to = "pop",
    names_transform = ~str_remove(., "x") %>% as.numeric, 
    values_transform = ~ as.numeric(.) * 1000
  ) %>% 
  filter(between(age, 13, 100)) %>% 
  mutate(
    age = cut(age, c(0, 34.5, 1000), labels = c("13-34", "35+"))
  )
pop <- pop %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(country, year, age) %>% 
  summarise(pop = sum(pop)) %>% 
  ungroup()
```

@@

```{r gallup-save}
# Save
dir.create("data", FALSE)
write_rds(gwp, "data-raw/gwp.rds")
write_rds(gbd, "data/gbd.rds")
write_rds(pop, "data/population.rds")
```

### 2022-11-09 ms.Rmd

keywords: well-being, social media
wordcount: "`r wordcountaddin:::word_count(filename = 'ms.Rmd')`"
bibliography: "references.bib"
floatsintext: yes
linenumbers: yes
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
header-includes:
  - |
    \setlength{\parskip}{0pt}
    \setlength\parindent{24pt}
csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: apa7
classoption: doc
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
always_allow_html: true
---

@@

We studied two metrics of Facebook adoption, daily (DAU) and monthly active users (MAU), from 2008 to 2019 for 72 countries (see Figure \@ref(fig:fig-forest)). FB estimates yearly DAU & MAU with their corresponding means from June 1 to August 31, and rounds values greater than 10,000,000 to three significant digits, and values less than 10,000,000 to the nearest 10,000. We report analyses based on DAU in the main manuscript, and MAU in the online supplement. Facebook calculates DAU and MAU estimates separately for individuals aged 13-34 and 35+. User age is determined based on Facebook profile information, which can be inaccurate (e.g. young users reporting an older age.) Accordingly, Facebook has trimmed 0.008% of total MAU to exclude accounts with unrealistic or non-reported ages.

@@

We examined the association between Facebook adoption and well-being through bayesian hierarchical regression models, estimated separately for DAU and MAU, and each of the three well-being outcomes. We regressed the outcome $y$ on time (decades, centered on 2014), within-country centered DAU or MAU (in separate models), age, sex, all the interactions except time by DAU/MAU, and the between-country centered DAU/MAU. We allowed all coefficients to vary randomly across the 72 countries, except the between-country predictor. Because we aggregated the outcomes, we included to standard errors of the outcome in the model to account for the varying group sizes and resulting uncertainties in the modelled data. Formally, we specified the model as

@@

# Average within- and between-country associations
tab_world <- coefs %>% 
  filter(
    country == "World",
    Hypothesis %in% c("Average", "Between-country")
  ) %>% 
  mutate(
    outcome = case_when(
      outcome == "Life_satisfaction" ~ "LS",
      outcome == "Negative_experiences" ~ "NE",
      outcome == "Positive_experiences" ~ "PE",
      TRUE ~ outcome
    )
  ) %>%
  mutate(
    across(c(Mean, CI_low, CI_high), list("str" = ~number(., .01))),
    result = str_glue(
      "{Mean_str} [{CI_low_str}, {CI_high_str}] ({percent2(pd)})"
    )
  ) %>% 
  select(
    predictor, Outcome = outcome, Hypothesis, result
  ) %>% 
  pivot_wider(names_from = Hypothesis, values_from = result) %>% 
  rename(
    `Within-country` = Average,
    `Between countries` = `Between-country`
  ) %>% 
  select(predictor, Outcome, `Between countries`, `Within-country`) %>% 
  arrange(predictor, Outcome)

tab_world %>% 
  filter(Outcome %in% c("LS", "NE", "PE")) %>% 
  mutate(
    Predictor = str_to_upper(predictor), 
    Predictor = if_else(1:n() %in% c(1, 4), Predictor, ""),
    .keep = "unused", .before = 1
  ) %>% 
  rename(X = Predictor, Y = Outcome) %>% 
  mutate(
    across(
      ends_with("country"), 
      ~if_else(str_starts(., "-"), ., paste0("\\phantom{-}", .))
    ),
    across(
      ends_with("country"),
      ~str_replace_all(., "%", "\\\\%")
    )
  ) %>% 
  apa_table(
    caption = "Average Facebook adoption and well-being associations",
    font_size = "scriptsize",
    note = "LS: Life satisfaction, NE: Negative experiences: PE: Positive experiences. Numbers indicate posterior means, [95\\%CIs], and (posterior probabilities of direction).",
    span_text_columns = FALSE,
    escape = FALSE,
    midrules = 3,
    placement = "htb",
    align = 'llll'
  )
```

Variations in such confounding factors are likely to be significantly smaller within countries but over time in a 12-year period. We therefore next focused on the model's within-country associations ($\alpha_4$). They measure the extent to which Facebook adoption in a given country predicted well-being in that country, adjusting for temporal trends in the country's well-being outcome. For the average country and across age and sex, we found that within-country increases in DAU predicted greater levels of life satisfaction and positive experiences, and lower levels of negative experiences, although only the latter association's 95%CI excluded zero (Table \@ref(tab:tab-world); Within-country).

However, these results are qualified because a focus on daily active users could miss those who use Facebook less regularly. To test this possibility, we also conducted our analyses using monthly active Facebook users (MAU) as the predictor. Both the between- and within-country associations linking MAU and well-being were very similar to those linking DAU and well-being, although in general of smaller magnitude, and the within-country link between MAU and negative experiences was not credibly different from zero.

@@

tmp %>%
  ggplot() +
  aes(
    x = Mean, 
    y = fct_rev(country), 
    col = Age, 
    fill = interaction(Age, cred, sep = " "), 
    shape = Sex
  ) +
  scale_y_discrete(
    "Country",
    expand = expansion(.015),
    labels = ~str_replace(., "World", "Average")
  ) +
  scale_x_continuous(
    str_glue(
      "Association with per capita Facebook users (95%CI)")
    ,
    breaks = extended_breaks(7),
    expand = expansion(.1)
  ) +
  scale_shape_manual(
    values = c(21, 22)
  ) +
  scale_color_brewer(
    "Age", palette = "Set1", aesthetics = c("color")
  ) +
  scale_fill_manual(values = c("white", "white", "#E41A1C", "#377EB8")) +
  geom_vline(xintercept = 0, lty = 2, size = .25, col = "grey50") +
  geom_linerangeh(
    aes(xmin = CI_low, xmax = CI_high),
    size = .15, position = position_dodge2v(.7), 
    show.legend = FALSE
  ) +
  geom_point(
    size = .65, stroke = .45,
    position = position_dodge2v(.7)
  ) +
  facet_nested(
    ~ predictor + outcome, scales = "free_x",
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(size = 1.5)),
    shape = guide_legend(override.aes = list(size = 1.5))
  ) +
  theme(
    text = element_text(size = 9),
    axis.text.y = element_text(size = rel(0.9), vjust = 0.35),
    axis.text.x = element_text(size = rel(0.8)),
    axis.title.y = element_blank(),
    panel.grid.major = element_line(
      size = .1, linetype = 2, color = "gray60"
    ),
    legend.position = "bottom", 
    legend.title = element_blank(),
    legend.box.margin = margin(-10, 0, -5, 0, "pt")
  )
```

@@

```{r gbd1}
#| include: !expr knitr::is_latex_output()
tab_world %>% 
  filter(Outcome %in% oo[4:6]) %>% 
  mutate(
    Predictor = str_to_upper(predictor), 
    Predictor = if_else(1:n() %in% c(1, 4), Predictor, ""),
    .keep = "unused", .before = 1
  ) %>% 
  rename(X = Predictor, Y = Outcome) %>% 
  mutate(
    across(
      ends_with("country"), 
      ~if_else(str_starts(., "-"), ., paste0("\\phantom{-}", .))
    ),
    across(
      ends_with("country"),
      ~str_replace_all(., "%", "\\\\%")
    )
  ) %>% 
  apa_table(
    caption = "Average Facebook adoption and mental health associations",
    font_size = "scriptsize",
    note = "Numbers indicate posterior means, [95\\%CIs], and (posterior probabilities of direction).",
    span_text_columns = FALSE,
    escape = FALSE,
    midrules = 3,
    placement = "htb",
    align = 'llll'
  )
```

>>>>

keywords: well-being, social media
wordcount: "`r wordcountaddin:::word_count(filename = 'ms.Rmd')`"
bibliography: "references.bib"
floatsintext: yesnc
linenumbers: yes
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
header-includes:
  - |
    \setlength{\parskip}{0pt}
    \setlength\parindent{24pt}
csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: apa7
classoption: doc
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
always_allow_html: true
---

@@

We studied two metrics of Facebook adoption, daily (DAU) and monthly active users (MAU), from 2008 to 2019 for 72 countries (see Figure \@ref(fig:fig-forest)). FB estimates yearly DAU & MAU with their corresponding means from June 1 to August 31, and rounds values greater than 10,000,000 to three significant digits, and values less than 10,000,000 to the nearest 10,000. Facebook calculates DAU and MAU estimates separately for individuals aged 13-34 and 35+. User age is determined based on Facebook profile information, which can be inaccurate (e.g. young users reporting an older age.) Accordingly, Facebook has trimmed 0.008% of total MAU to exclude accounts with unrealistic or non-reported ages.

@@

We examined the association between Facebook adoption and well-being through bayesian hierarchical regression models, estimated separately for DAU and MAU, and each of the three well-being outcomes. We regressed the outcome $y$ on time (decades, centered on 2014), within-country centered DAU or MAU (in separate models), age, sex, all the interactions except time by DAU/MAU, and the between-country centered DAU/MAU. We allowed all coefficients to vary randomly across the 72 countries, except the between-country predictor. Because we aggregated the outcomes, we included standard errors of the outcome in the model to account for the varying group sizes and resulting uncertainties in the modelled data. Formally, we specified the model as

@@

# Average within- and between-country associations
tab_world <- coefs %>% 
  filter(
    country == "World",
    Hypothesis %in% c("Average", "Between-country")
  ) %>% 
  mutate(
    across(c(Mean, CI_low, CI_high), list("str" = ~number(., .01))),
    result = str_glue(
      "{Mean_str} [{CI_low_str}, {CI_high_str}] ({percent2(pd)})"
    )
  ) %>% 
  select(
    predictor, Outcome = outcome, Hypothesis, result
  ) %>% 
  pivot_wider(names_from = Hypothesis, values_from = result) %>% 
  rename(
    `Within-country` = Average,
    `Between countries` = `Between-country`
  ) %>% 
  select(predictor, Outcome, `Between countries`, `Within-country`) %>% 
  arrange(predictor, Outcome)

tab_world %>% 
  filter(Outcome %in% oo[1:3]) %>% 
  mutate(
    Outcome = str_replace(Outcome, "_", " "),
    Predictor = str_to_upper(predictor), 
    Predictor = if_else(1:n() %in% c(1, 4), Predictor, ""),
    .keep = "unused", .before = 1
  ) %>% 
  mutate(
    across(
      ends_with("country"), 
      ~if_else(str_starts(., "-"), ., paste0("\\phantom{-}", .))
    ),
    across(
      ends_with("country"),
      ~str_replace_all(., "%", "\\\\%")
    )
  ) %>% 
  apa_table(
    caption = "Average Facebook adoption and well-being associations",
    font_size = "scriptsize",
    note = "LS: Life satisfaction, NE: Negative experiences: PE: Positive experiences. Numbers indicate posterior means, [95\\%CIs], and (posterior probabilities of direction).",
    span_text_columns = FALSE,
    escape = FALSE,
    midrules = 3,
    placement = "htb",
    align = 'llll'
  )
```

Variations in such confounding factors are likely to be significantly smaller within countries but over time in a 12-year period. We therefore next focused on the model's within-country associations ($\alpha_4$). They measure the extent to which Facebook adoption in a given country predicted well-being in that country, adjusting for temporal trends in the country's well-being outcome. For the average country and across age and sex, we found that within-country increases in DAU predicted greater levels of life satisfaction and positive experiences, and lower levels of negative experiences, although only the positive experience association's 95%CI excluded zero (Table \@ref(tab:tab-world); Within-country).

However, these results are qualified because a focus on daily active users could miss those who use Facebook less regularly. To test this possibility, we also conducted our analyses using monthly active Facebook users (MAU) as the predictor. Both the between- and within-country associations linking MAU and well-being were very similar to those linking DAU and well-being, although in general of smaller magnitude. Moreover, the within-country link between MAU and negative experiences' credibility interval was narrowly centered on zero, indicating relative certainty that the association is practically equivalent to zero.

@@

tmp %>%
  ggplot() +
  aes(
    x = Mean, 
    y = fct_rev(country), 
    col = Age, 
    fill = interaction(Age, cred, sep = " "), 
    shape = Sex
  ) +
  scale_y_discrete(
    "Country",
    expand = expansion(.015),
    labels = ~str_replace(., "World", "Average")
  ) +
  scale_x_continuous(
    str_glue(
      "Association with per capita Facebook users (95%CI)")
    ,
    breaks = extended_breaks(7),
    expand = expansion(.1)
  ) +
  scale_shape_manual(
    values = c(21, 22)
  ) +
  scale_color_brewer(
    "Age", palette = "Set1", aesthetics = c("color")
  ) +
  scale_fill_manual(values = c("white", "white", "#E41A1C", "#377EB8")) +
  geom_vline(xintercept = 0, lty = 2, linewidth = .25, col = "grey50") +
  geom_linerangeh(
    aes(xmin = CI_low, xmax = CI_high),
    size = .15, position = position_dodge2v(.7), 
    show.legend = FALSE
  ) +
  geom_point(
    size = .65, stroke = .45,
    position = position_dodge2v(.7)
  ) +
  facet_nested(
    ~ predictor + outcome, scales = "free_x",
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(size = 1.5)),
    shape = guide_legend(override.aes = list(size = 1.5))
  ) +
  theme(
    text = element_text(size = 9),
    axis.text.y = element_text(size = rel(0.9), vjust = 0.35),
    axis.text.x = element_text(size = rel(0.8)),
    axis.title.y = element_blank(),
    panel.grid.major = element_line(
      size = .1, linetype = 2, color = "gray60"
    ),
    legend.position = "bottom", 
    legend.title = element_blank(),
    legend.box.margin = margin(-10, 0, -5, 0, "pt")
  )
```

@@

```{r gbd1}
#| include: !expr knitr::is_latex_output()
tab_world %>% 
  filter(Outcome %in% oo[4:6]) %>% 
  mutate(
    Predictor = str_to_upper(predictor), 
    Predictor = if_else(1:n() %in% c(1, 4), Predictor, ""),
    .keep = "unused", .before = 1
  ) %>% 
  mutate(
    across(
      ends_with("country"), 
      ~if_else(str_starts(., "-"), ., paste0("\\phantom{-}", .))
    ),
    across(
      ends_with("country"),
      ~str_replace_all(., "%", "\\\\%")
    )
  ) %>% 
  apa_table(
    caption = "Average Facebook adoption and mental health associations",
    font_size = "scriptsize",
    note = "Numbers indicate posterior means, [95\\%CIs], and (posterior probabilities of direction).",
    span_text_columns = FALSE,
    escape = FALSE,
    midrules = 3,
    placement = "htb",
    align = 'llll'
  )
```
